---
title: "Zanzibar red colobus helminth manuscript analysis"
output: html_document
---

# Overview

Analysis code for the manuscript *Gastrointestinal helminth infections in the Zanzibar red colobus monkeys (Piliocolobus kirkii)*. The workflow starts from the `All_Samples` worksheet in the master dataset and reproduces the analyses and figures reported in the manuscript and supplementary material.

# 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lme4)
library(emmeans)
library(DescTools)
library(rstatix)
library(ggsignif)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
```

```{r load-data}
data_raw <- read_excel("ZRC_helminth_data.xlsx", sheet = "All_Samples")
```

# 2. Data preparation

```{r prepare-data}
data <- data_raw %>%
  mutate(
    parasite_taxon = na_if(trimws(parasite_taxon), ""),
    year = factor(year),
    protocol = factor(protocol),
    method = factor(method, levels = c("DS", "FL", "SD")),
    slide = as.integer(slide),
    ID = trimws(ID),
    ID = recode(ID,
      `Droopy lip` = "Droopy Lip",
      `Tight lip` = "Tight Lip"
    ),
    known_ID = ID != "UNK",
    individual_ID = if_else(
      known_ID,
      paste(group, ID, sep = "_"),
      NA_character_
    ),
    individual_ID = factor(individual_ID),
    detected = as.integer(!is.na(parasite_taxon))
  )

main_data <- data %>%
  filter(year %in% c("2023", "2024"))

sample_data <- main_data %>%
  group_by(
    code, sample, group, ID, known_ID, individual_ID,
    lab, year, protocol, storage_time
  ) %>%
  summarise(
    positive = as.integer(any(!is.na(parasite_taxon))),
    richness = n_distinct(parasite_taxon[!is.na(parasite_taxon)]),
    .groups = "drop"
  )

single_slide <- main_data %>%
  filter(slide == 1)

single_slide_samples <- single_slide %>%
  group_by(
    code, sample, group, ID, known_ID, individual_ID,
    lab, year, protocol, storage_time
  ) %>%
  summarise(
    positive = as.integer(any(!is.na(parasite_taxon))),
    richness = n_distinct(parasite_taxon[!is.na(parasite_taxon)]),
    .groups = "drop"
  )

method_single_slide <- single_slide %>%
  group_by(
    code, sample, group, ID, known_ID, individual_ID,
    lab, year, protocol, method
  ) %>%
  summarise(
    positive = as.integer(any(!is.na(parasite_taxon))),
    .groups = "drop"
  )

data_2024 <- data %>%
  filter(year == "2024", protocol == "centrifugal")

cumulative_screening_effort <- data_2024 %>%
  crossing(slides_examined = 1:4) %>%
  filter(slide <= slides_examined) %>%
  group_by(code, slides_examined) %>%
  summarise(
    positive = as.integer(any(!is.na(parasite_taxon))),
    richness = n_distinct(parasite_taxon[!is.na(parasite_taxon)]),
    .groups = "drop"
  )

data_2025 <- data %>%
  filter(year == "2025")

protocol_2025_samples <- data_2025 %>%
  group_by(sample, ID, individual_ID, group, protocol) %>%
  summarise(
    positive = as.integer(any(!is.na(parasite_taxon))),
    richness = n_distinct(parasite_taxon[!is.na(parasite_taxon)]),
    .groups = "drop"
  )

protocol_2025_wide <- protocol_2025_samples %>%
  pivot_wider(
    names_from = protocol,
    values_from = c(positive, richness)
  )
```

# 3. Descriptive results

## Sample positivity and richness

```{r overall-descriptives}
overall_summary <- sample_data %>%
  group_by(year) %>%
  summarise(
    n_samples = n(),
    n_positive = sum(positive),
    percent_positive = 100 * mean(positive),
    mean_richness = mean(richness),
    sd_richness = sd(richness),
    .groups = "drop"
  )

overall_summary

sample_data %>%
  summarise(
    n_samples = n(),
    n_positive = sum(positive),
    percent_positive = 100 * mean(positive),
    mean_richness = mean(richness),
    sd_richness = sd(richness),
    min_richness = min(richness),
    max_richness = max(richness)
  )
```

## Taxon detection

```{r taxon-detection}
sample_taxa <- main_data %>%
  filter(!is.na(parasite_taxon)) %>%
  distinct(code, year, parasite_taxon)

taxon_detection_overall <- sample_taxa %>%
  count(parasite_taxon, name = "n_positive") %>%
  mutate(
    n_samples = n_distinct(main_data$code),
    percent_positive = 100 * n_positive / n_samples
  ) %>%
  arrange(desc(percent_positive))

taxon_detection_by_year <- sample_taxa %>%
  count(year, parasite_taxon, name = "n_positive") %>%
  left_join(sample_data %>% count(year, name = "n_samples"), by = "year") %>%
  mutate(percent_positive = 100 * n_positive / n_samples) %>%
  arrange(year, desc(percent_positive))

taxon_detection_overall
taxon_detection_by_year
```

## Co-infection

```{r coinfection-descriptives}
coinfection_full <- sample_data %>%
  summarise(
    n_samples = n(),
    n_positive = sum(richness > 0),
    n_coinfected = sum(richness > 1),
    percent_all_samples = 100 * n_coinfected / n_samples,
    percent_positive_samples = 100 * n_coinfected / n_positive
  )

coinfection_full
```

## Identified individuals and repeated sampling

```{r sampling-structure}
sampling_structure <- sample_data %>%
  group_by(year) %>%
  summarise(
    total_samples = n(),
    identified_samples = sum(known_ID),
    percent_identified = 100 * identified_samples / total_samples,
    unique_individuals = n_distinct(individual_ID[known_ID]),
    .groups = "drop"
  )

samples_per_individual <- sample_data %>%
  filter(known_ID) %>%
  count(year, individual_ID, name = "n_samples")

repeated_sampling_summary <- samples_per_individual %>%
  group_by(year) %>%
  summarise(
    unique_individuals = n(),
    mean_samples = mean(n_samples),
    median_samples = median(n_samples),
    min_samples = min(n_samples),
    max_samples = max(n_samples),
    .groups = "drop"
  )

sampling_structure
repeated_sampling_summary
```

# 4. Standardised 2023–2024 comparison

The between-period analyses use one slide per coproscopic technique in both sampling periods.

## Overall sample positivity

```{r positivity-glmm}
positivity_model <- glmer(
  positive ~ year + (1 | group),
  data = single_slide_samples,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

summary(positivity_model)
```

## Helminth richness

```{r richness-glmm}
richness_model <- glmer(
  richness ~ year + (1 | group),
  data = single_slide_samples,
  family = poisson,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

summary(richness_model)
```

## Co-infection

```{r coinfection-glmm}
coinfection_data <- single_slide_samples %>%
  mutate(coinfected = as.integer(richness >= 2))

coinfection_summary <- coinfection_data %>%
  group_by(year) %>%
  summarise(
    n_samples = n(),
    n_coinfected = sum(coinfected),
    percent_coinfected = 100 * mean(coinfected),
    .groups = "drop"
  )

coinfection_model <- glmer(
  coinfected ~ year + (1 | group),
  data = coinfection_data,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

coinfection_null <- glmer(
  coinfected ~ 1 + (1 | group),
  data = coinfection_data,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

coinfection_summary
anova(coinfection_null, coinfection_model, test = "Chisq")
```

# 5. Coproscopic technique by sampling period

```{r technique-glmm}
detection_model_data <- method_single_slide %>%
  mutate(
    year = factor(year),
    method = factor(method, levels = c("DS", "FL", "SD")),
    group = factor(group),
    code = factor(code)
  )

detection_model_no_interaction <- glmer(
  positive ~ year + method + (1 | group) + (1 | code),
  data = detection_model_data,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

detection_model <- glmer(
  positive ~ year * method + (1 | group) + (1 | code),
  data = detection_model_data,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

anova(detection_model_no_interaction, detection_model, test = "Chisq")
summary(detection_model)
```

```{r technique-contrasts}
method_emm <- emmeans(detection_model, ~ method | year, type = "response")
pairs(method_emm, adjust = "tukey")

year_emm <- emmeans(detection_model, ~ year | method, type = "response")

year_contrasts <- contrast(year_emm, method = "pairwise", adjust = "none") %>%
  summary(type = "response") %>%
  as.data.frame() %>%
  mutate(p_holm = p.adjust(p.value, method = "holm"))

year_contrasts
```

# 6. Screening effort in 2024

## Sample positivity

```{r screening-positivity}
screening_summary <- cumulative_screening_effort %>%
  group_by(slides_examined) %>%
  summarise(
    n_samples = n(),
    n_positive = sum(positive),
    percent_positive = 100 * mean(positive),
    mean_richness = mean(richness),
    .groups = "drop"
  )

screening_positive_wide <- cumulative_screening_effort %>%
  select(code, slides_examined, positive) %>%
  pivot_wider(
    names_from = slides_examined,
    values_from = positive,
    names_prefix = "slides_"
  )

cochran_result <- CochranQTest(
  as.matrix(screening_positive_wide %>% select(slides_1:slides_4))
)

screening_summary
cochran_result
```

```{r screening-positivity-pairwise}
exact_screening_comparison <- function(data, slide_a, slide_b) {
  wide <- data %>%
    filter(slides_examined %in% c(slide_a, slide_b)) %>%
    select(code, slides_examined, positive) %>%
    pivot_wider(
      names_from = slides_examined,
      values_from = positive,
      names_prefix = "slide_"
    )

  a <- wide[[paste0("slide_", slide_a)]]
  b <- wide[[paste0("slide_", slide_b)]]

  lost <- sum(a == 1 & b == 0)
  gained <- sum(a == 0 & b == 1)
  discordant <- lost + gained

  p <- if (discordant > 0) {
    binom.test(gained, discordant, p = 0.5)$p.value
  } else {
    1
  }

  tibble(
    comparison = paste(slide_a, "vs", slide_b),
    gained_detection = gained,
    lost_detection = lost,
    discordant_pairs = discordant,
    p_value = p
  )
}

screening_pairwise <- bind_rows(
  exact_screening_comparison(cumulative_screening_effort, 1, 2),
  exact_screening_comparison(cumulative_screening_effort, 2, 3),
  exact_screening_comparison(cumulative_screening_effort, 3, 4),
  exact_screening_comparison(cumulative_screening_effort, 2, 4),
  exact_screening_comparison(cumulative_screening_effort, 1, 4)
) %>%
  mutate(p_holm = p.adjust(p_value, method = "holm"))

screening_pairwise
```

## Richness

```{r screening-richness}
screening_richness_wide <- cumulative_screening_effort %>%
  select(code, slides_examined, richness) %>%
  pivot_wider(
    names_from = slides_examined,
    values_from = richness,
    names_prefix = "slides_"
  )

friedman_result <- friedman.test(
  as.matrix(screening_richness_wide %>% select(slides_1:slides_4))
)

richness_comparisons <- list(
  "1 vs 2" = c("slides_1", "slides_2"),
  "2 vs 3" = c("slides_2", "slides_3"),
  "3 vs 4" = c("slides_3", "slides_4"),
  "2 vs 4" = c("slides_2", "slides_4"),
  "1 vs 4" = c("slides_1", "slides_4")
)

richness_pairwise <- map_dfr(names(richness_comparisons), function(comp) {
  vars <- richness_comparisons[[comp]]
  test <- wilcox.test(
    screening_richness_wide[[vars[1]]],
    screening_richness_wide[[vars[2]]],
    paired = TRUE,
    exact = FALSE
  )

  tibble(
    comparison = comp,
    mean_first = mean(screening_richness_wide[[vars[1]]]),
    mean_second = mean(screening_richness_wide[[vars[2]]]),
    V = unname(test$statistic),
    p_value = test$p.value
  )
}) %>%
  mutate(p_holm = p.adjust(p_value, method = "holm"))

richness_gains <- screening_richness_wide %>%
  summarise(
    gained_1_to_2 = sum(slides_2 > slides_1),
    gained_2_to_3 = sum(slides_3 > slides_2),
    gained_3_to_4 = sum(slides_4 > slides_3),
    gained_2_to_4 = sum(slides_4 > slides_2)
  )

friedman_result
richness_pairwise
richness_gains
```

# 7. Individuals sampled in both periods

All standardised faecal samples from the 15 identified individuals sampled in both 2023 and 2024 are retained.

```{r matched-data}
IDs_both_years <- single_slide_samples %>%
  filter(known_ID) %>%
  distinct(individual_ID, year) %>%
  count(individual_ID, name = "n_years") %>%
  filter(n_years == 2) %>%
  pull(individual_ID)

matched_repeated <- single_slide_samples %>%
  filter(known_ID, individual_ID %in% IDs_both_years) %>%
  droplevels()

matched_repeated %>%
  summarise(
    n_samples = n(),
    n_individuals = n_distinct(individual_ID)
  )

matched_repeated %>% count(year)
```

```{r matched-positivity}
matched_positive_model <- glmer(
  positive ~ year + (1 | individual_ID),
  data = matched_repeated,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

matched_positive_null <- glmer(
  positive ~ 1 + (1 | individual_ID),
  data = matched_repeated,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

matched_repeated %>%
  group_by(year) %>%
  summarise(
    n_samples = n(),
    n_positive = sum(positive),
    percent_positive = 100 * mean(positive),
    .groups = "drop"
  )

anova(matched_positive_null, matched_positive_model, test = "Chisq")
```

```{r matched-richness}
matched_richness_model <- glmer(
  richness ~ year + (1 | individual_ID),
  data = matched_repeated,
  family = poisson,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

matched_richness_null <- glmer(
  richness ~ 1 + (1 | individual_ID),
  data = matched_repeated,
  family = poisson,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

matched_repeated %>%
  group_by(year) %>%
  summarise(
    n_samples = n(),
    mean_richness = mean(richness),
    sd_richness = sd(richness),
    .groups = "drop"
  )

anova(matched_richness_null, matched_richness_model, test = "Chisq")
```

# 8. 2025 paired protocol comparison

```{r protocol-2025-descriptives}
protocol_2025_samples %>%
  group_by(protocol) %>%
  summarise(
    n = n(),
    n_positive = sum(positive),
    percent_positive = 100 * mean(positive),
    mean_richness = mean(richness),
    sd_richness = sd(richness),
    .groups = "drop"
  )
```

```{r protocol-2025-tests}
tab_2025 <- table(
  protocol_2025_wide$positive_fieldfriendly,
  protocol_2025_wide$positive_centrifugal
)

fieldfriendly_only <- tab_2025["1", "0"]
centrifugal_only <- tab_2025["0", "1"]

exact_mcnemar_2025 <- binom.test(
  x = fieldfriendly_only,
  n = fieldfriendly_only + centrifugal_only,
  p = 0.5,
  alternative = "two.sided"
)

wilcox_2025 <- wilcox.test(
  protocol_2025_wide$richness_fieldfriendly,
  protocol_2025_wide$richness_centrifugal,
  paired = TRUE,
  exact = FALSE
)

tab_2025
exact_mcnemar_2025
wilcox_2025
```

# 9. Laboratory sensitivity analyses

```{r laboratory-sensitivity}
lab_2024_samples <- single_slide_samples %>%
  filter(year == "2024")

lab_positivity_model <- glmer(
  positive ~ lab + (1 | group),
  data = lab_2024_samples,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

lab_richness_model <- glmer(
  richness ~ lab + (1 | group),
  data = lab_2024_samples,
  family = poisson,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 100000)
  )
)

summary(lab_positivity_model)
summary(lab_richness_model)
```

# 10. Storage-time sensitivity analyses

These models provide the results reported in Supplementary Table S2.

```{r storage-time}
storage_2023_2024 <- sample_data %>%
  select(year, positive, richness, storage_time) %>%
  mutate(storage_time = as.numeric(storage_time))

storage_2025 <- protocol_2025_samples %>%
  left_join(
    data_2025 %>% distinct(sample, protocol, storage_time),
    by = c("sample", "protocol")
  ) %>%
  transmute(
    year = factor("2025", levels = levels(data$year)),
    positive,
    richness,
    storage_time = as.numeric(storage_time)
  )

storage_data <- bind_rows(storage_2023_2024, storage_2025)

storage_models <- list(
  positivity_2023 = glm(
    positive ~ storage_time,
    data = filter(storage_data, year == "2023"),
    family = binomial
  ),
  positivity_2024 = glm(
    positive ~ storage_time,
    data = filter(storage_data, year == "2024"),
    family = binomial
  ),
  positivity_2025 = glm(
    positive ~ storage_time,
    data = filter(storage_data, year == "2025"),
    family = binomial
  ),
  richness_2023 = glm(
    richness ~ storage_time,
    data = filter(storage_data, year == "2023"),
    family = poisson
  ),
  richness_2024 = glm(
    richness ~ storage_time,
    data = filter(storage_data, year == "2024"),
    family = poisson
  ),
  richness_2025 = glm(
    richness ~ storage_time,
    data = filter(storage_data, year == "2025"),
    family = poisson
  )
)

map(storage_models, summary)
```

# 11. Main figures

## Figure 1. Protocol comparison

```{r figure-1-protocol}
protocol_comparison <- grViz("
digraph protocol_comparison {

  graph [
    layout = dot,
    rankdir = TB,
    bgcolor = 'transparent',
    splines = ortho,
    nodesep = 0.35,
    ranksep = 0.55,
    margin = 0.2,
    pad = 0.2
  ]

  node [
    shape = box,
    style = 'filled',
    fontname = Helvetica,
    fontsize = 15,
    color = 'black',
    penwidth = 1.2,
    margin = 0.18
  ]

  edge [
    color = 'black',
    penwidth = 1.2,
    arrowsize = 0.7
  ]

  protocol23 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><FONT POINT-SIZE='18'><B>2023 FIELD-FRIENDLY PROTOCOL</B></FONT></TD></TR>
        <TR><TD HEIGHT='8'></TD></TR>
        <TR><TD ALIGN='CENTER'>Partial homogenisation</TD></TR>
        <TR><TD ALIGN='CENTER'>27–30% unbuffered formalin</TD></TR>
        <TR><TD ALIGN='CENTER'>1 slide per technique</TD></TR>
      </TABLE>
    >,
    fillcolor = 'lightgrey'
  ]

  ds23 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>Direct smear</B></TD></TR>
        <TR><TD ALIGN='CENTER'>Faecal–formalin</TD></TR>
        <TR><TD ALIGN='CENTER'>suspension</TD></TR>
      </TABLE>
    >,
    fillcolor = 'white', width = 2.2, height = 0.9, fixedsize = true
  ]

  fl23 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>Flotation</B></TD></TR>
        <TR><TD ALIGN='CENTER'>Saturated NaCl</TD></TR>
        <TR><TD ALIGN='CENTER'>15 min</TD></TR>
      </TABLE>
    >,
    fillcolor = 'white', width = 2.2, height = 0.9, fixedsize = true
  ]

  sd23 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>Sedimentation</B></TD></TR>
        <TR><TD ALIGN='CENTER'>Water wash ×3</TD></TR>
      </TABLE>
    >,
    fillcolor = 'white', width = 2.2, height = 0.9, fixedsize = true
  ]

  protocol24 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><FONT POINT-SIZE='18'><B>2024 CENTRIFUGAL PROTOCOL</B></FONT></TD></TR>
        <TR><TD HEIGHT='8'></TD></TR>
        <TR><TD ALIGN='CENTER'>Thorough homogenisation</TD></TR>
        <TR><TD ALIGN='CENTER'>10% buffered formalin + 0.03% Eosin</TD></TR>
        <TR><TD ALIGN='CENTER'>4 slides per technique</TD></TR>
      </TABLE>
    >,
    fillcolor = 'lightgrey'
  ]

  processing24 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>CENTRIFUGAL PREPROCESSING</B></TD></TR>
        <TR><TD ALIGN='CENTER'>Fine-mesh sieve</TD></TR>
        <TR><TD ALIGN='CENTER'>Rinsed with water to 35 ml</TD></TR>
        <TR><TD ALIGN='CENTER'>2000 rpm, 5 min</TD></TR>
        <TR><TD ALIGN='CENTER'>Faecal sediment retained</TD></TR>
      </TABLE>
    >,
    fillcolor = 'grey95'
  ]

  ds24 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>Direct smear</B></TD></TR>
        <TR><TD ALIGN='CENTER'>Sieved faecal material</TD></TR>
        <TR><TD ALIGN='CENTER'>+ water</TD></TR>
      </TABLE>
    >,
    fillcolor = 'white', width = 2.2, height = 0.9, fixedsize = true
  ]

  fl24 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>Flotation</B></TD></TR>
        <TR><TD ALIGN='CENTER'>Saturated NaCl</TD></TR>
        <TR><TD ALIGN='CENTER'>2000 rpm, 5 min</TD></TR>
      </TABLE>
    >,
    fillcolor = 'white', width = 2.2, height = 0.9, fixedsize = true
  ]

  sd24 [
    label = <
      <TABLE BORDER='0' CELLBORDER='0' CELLSPACING='0' CELLPADDING='0'>
        <TR><TD ALIGN='CENTER'><B>Sedimentation</B></TD></TR>
        <TR><TD ALIGN='CENTER'>0.9% NaCl wash ×4</TD></TR>
      </TABLE>
    >,
    fillcolor = 'white', width = 2.2, height = 0.9, fixedsize = true
  ]

  protocol23 -> ds23
  protocol23 -> fl23
  protocol23 -> sd23
  { rank = same; ds23; fl23; sd23 }

  protocol24 -> processing24
  processing24 -> ds24
  processing24 -> fl24
  processing24 -> sd24 [style = invis]
  fl24 -> sd24
  { rank = same; ds24; fl24; sd24 }

  { rank = same; protocol23; protocol24 }
  protocol23 -> protocol24 [style = invis, weight = 20]
}
")

protocol_comparison
```

## Figure 2. Cumulative sample positivity

```{r figure-2}
figure_2 <- ggplot(
  screening_summary,
  aes(x = factor(slides_examined), y = percent_positive)
) +
  geom_col(fill = "darkgrey", colour = "black", width = 0.6) +
  labs(
    x = "Number of slides examined",
    y = "Helminth-positive samples (%)"
  ) +
  theme_bw(base_size = 16) +
  coord_cartesian(ylim = c(0, 70))

figure_2
```

## Figure 3. Cumulative richness

```{r figure-3}
richness_plot_df <- cumulative_screening_effort %>%
  mutate(
    slide_label = factor(slides_examined, levels = 1:4),
    richness_group = factor(richness, levels = 0:3)
  )

figure_3 <- ggplot(
  richness_plot_df,
  aes(x = slide_label, fill = richness_group)
) +
  geom_bar(position = "fill", width = 0.65, colour = "black") +
  scale_y_continuous(
    breaks = seq(0, 1, 0.2),
    labels = scales::label_percent(),
    limits = c(0, 1.04),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_fill_manual(
    values = c(
      "0" = "grey90",
      "1" = "grey70",
      "2" = "grey45",
      "3" = "grey20"
    )
  ) +
  labs(
    x = "Number of slides examined",
    y = "Samples (%)",
    fill = "Helminth\nrichness"
  ) +
  theme_bw(base_size = 16)

figure_3
```

## Figure 4. Technique-specific sample positivity

```{r figure-4-data}
method_labels <- c(
  DS = "Direct smear",
  FL = "Flotation",
  SD = "Sedimentation"
)

figure_4_data <- method_single_slide %>%
  group_by(year, method) %>%
  summarise(
    n_samples = n(),
    n_positive = sum(positive),
    percent_positive = 100 * mean(positive),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    ci_lower = 100 * binom.test(n_positive, n_samples)$conf.int[1],
    ci_upper = 100 * binom.test(n_positive, n_samples)$conf.int[2]
  ) %>%
  ungroup() %>%
  mutate(
    year = factor(as.character(year), levels = c("2023", "2024")),
    method = recode(as.character(method), !!!method_labels)
  )

figure_4_annotation <- tibble(
  method = "Flotation",
  year1 = "2023",
  year2 = "2024",
  y_position = 55,
  annotations = "***"
)
```

```{r figure-4}
figure_4 <- ggplot(
  figure_4_data,
  aes(x = year, y = percent_positive, fill = year)
) +
  geom_col(width = 0.6, colour = "black") +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    width = 0.15
  ) +
  scale_fill_grey(start = 0.8, end = 0.4, guide = "none") +
  facet_wrap(~method, nrow = 1) +
  coord_cartesian(ylim = c(0, 60)) +
  scale_y_continuous(
    breaks = seq(0, 60, by = 10),
    minor_breaks = seq(0, 60, by = 5)
  ) +
  labs(
    x = "Year of sampling",
    y = "Helminth-positive samples (%)"
  ) +
  theme_bw(base_size = 16) +
  theme(panel.grid.major.x = element_blank()) +
  geom_signif(
    data = figure_4_annotation,
    aes(
      xmin = year1,
      xmax = year2,
      annotations = annotations,
      y_position = y_position
    ),
    manual = TRUE,
    tip_length = 0.01,
    inherit.aes = FALSE
  )

figure_4
```

# 12. Supplementary figures

## Figure S1. Technique-specific cumulative sample positivity

```{r figure-s1}
method_slide_detection <- main_data %>%
  filter(year == "2024") %>%
  mutate(
    method = factor(
      method,
      levels = c("DS", "FL", "SD"),
      labels = c("Direct smear", "Flotation", "Sedimentation")
    )
  ) %>%
  crossing(slides_examined = 1:4) %>%
  filter(slide <= slides_examined) %>%
  group_by(code, method, slides_examined) %>%
  summarise(
    positive = any(!is.na(parasite_taxon)),
    .groups = "drop"
  ) %>%
  group_by(method, slides_examined) %>%
  summarise(
    n_samples = n(),
    n_positive = sum(positive),
    percent_positive = 100 * n_positive / n_samples,
    .groups = "drop"
  )

figure_S1 <- ggplot(
  method_slide_detection,
  aes(
    x = slides_examined,
    y = percent_positive,
    linetype = method,
    group = method
  )
) +
  geom_line(linewidth = 0.8, colour = "black") +
  geom_point(size = 2.5, colour = "black") +
  scale_x_continuous(breaks = 1:4) +
  scale_y_continuous(
    breaks = seq(0, 50, by = 10),
    limits = c(0, 50),
    expand = expansion(mult = c(0, 0))
  ) +
  labs(
    x = "Number of slides examined",
    y = "Helminth-positive samples (%)",
    linetype = "Method"
  ) +
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    legend.position = "right"
  )

figure_S1
```

## Figure S2. Technique-specific cumulative richness

```{r figure-s2}
method_slide_richness <- main_data %>%
  filter(year == "2024") %>%
  mutate(
    method = factor(
      method,
      levels = c("DS", "FL", "SD"),
      labels = c("Direct smear", "Flotation", "Sedimentation")
    )
  ) %>%
  crossing(slides_examined = 1:4) %>%
  filter(slide <= slides_examined) %>%
  group_by(code, method, slides_examined) %>%
  summarise(
    richness = n_distinct(parasite_taxon[!is.na(parasite_taxon)]),
    .groups = "drop"
  ) %>%
  group_by(method, slides_examined) %>%
  summarise(
    mean_richness = mean(richness),
    .groups = "drop"
  )

figure_S2 <- ggplot(
  method_slide_richness,
  aes(
    x = slides_examined,
    y = mean_richness,
    linetype = method,
    group = method
  )
) +
  geom_line(linewidth = 0.8, colour = "black") +
  geom_point(size = 2.5, colour = "black") +
  scale_x_continuous(breaks = 1:4) +
  scale_y_continuous(
    breaks = seq(0, 1.0, by = 0.2),
    limits = c(0, 1.3),
    expand = expansion(mult = c(0, 0))
  ) +
  labs(
    x = "Number of slides examined",
    y = "Mean helminth richness\n(taxa per sample)",
    linetype = "Method"
  ) +
  theme_bw(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    legend.position = "right"
  )

figure_S2
```

# 13. Export figures

```{r export-figures, eval=FALSE}
ggsave("Figure_2.png", figure_2, width = 17, height = 12, units = "cm", dpi = 300)
ggsave("Figure_3.png", figure_3, width = 17, height = 12, units = "cm", dpi = 300)
ggsave("Figure_4.png", figure_4, width = 17, height = 12, units = "cm", dpi = 300)
ggsave("Figure_S1.png", figure_S1, width = 17, height = 12, units = "cm", dpi = 300)
ggsave("Figure_S2.png", figure_S2, width = 17, height = 12, units = "cm", dpi = 300)

svg_code <- DiagrammeRsvg::export_svg(protocol_comparison)
rsvg::rsvg_png(
  charToRaw(svg_code),
  file = "Figure_1_protocol_comparison.png",
  width = 8000
)
```

# 14. Reproducibility

```{r package-versions}
packages <- c(
  "tidyverse", "readxl", "lme4", "emmeans", "DescTools",
  "rstatix", "ggsignif", "DiagrammeR", "DiagrammeRsvg", "rsvg"
)

setNames(lapply(packages, packageVersion), packages)
```

```{r session-info}
sessionInfo()
```
